VIPRE: Virtualized Intermittently-Powered Runtime Environment

### 2017-07-09 paper/intro.tex
References and text up to paragraph 3 (intro) fixed

Advances in processor efficiency along with the development of energy-harvesting power systems has created a new category of devices that require neither a battery nor a tethered power supply~\cite{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}. These devices operate entirely using energy extracted from their environment, such as RF energy~\cite{rf_powered_computing_gollakota_2014}, photovoltaic components~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}. Incorporating compute, storage, sensing, and communication hardware, such energy-independent devices are a promising candidate as the underlying technology for emerging Internet of Things (IoT) applications~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical applications, and emerging applications, like energy-harvesting nano-satellites~\cite{kicksat}.

While promising, energy-harvesting devices present a unique design challenge in
that they operate only {\em intermittently} when energy is available. An
energy-harvesting device buffers energy from its environment, e.g., in a
capacitor. When a threshold quantity of energy accumulates, the device begins
operating. Harvestable energy sources typically produce relatively very little
power compared to a platform's operating power level. A devices operates for a
brief period of time and when buffered energy is depleted, shuts down and
begins recharging to later operate again. Charge and discharge times vary by
device and some devices~\cite{wisp} may fail 10 to 100 times per second~\cite{}.

Software in an energy-harvesting system operates according to the {\em
intermittent execution model}~\cite{dino}, with bursts of execution that are
interrupted by failure periods. When power fails, a device loses its volatile
state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e.,
FRAM, Flash. While capturing periodic checkpoints~\cite{mementos,tictpl,quickrecall}
and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve
execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\cite{mspcdino,edb,edbtoppicks}.

\TODO{Don't forget to include all the references in this section!}

%FRAM costs more than FRAM -- need to figure out how to execute out of SRAM
New programming and execution models~\cite{dino,ratchet,chain,alpaca} and
architecture mechanisms~\cite{clank,idetic,nvp} help avoid data
inconsistency. New architectures require costly hardware changes and are
inapplicable to today's systems. New programming models and compilers force
programmers to use new memory models~\cite{chain,ratchet}, which may inhibit
adoption. New memory models may also increase the amount of non-volatile
memory accesses by requiring non-volatile versioning~\cite{chain} or completely
precluding the use of volatile memory~\cite{ratchet}. Increasing pressure on
non-volatile memory or precluding the use of volatile memory may degrade
efficiency; e.g., FRAM has a higher access energy and latency than
SRAM~\cite{alpaca,wisp}. Commercially available devices are often provisioned
with much less SRAM than FRAM~\cite{wolverine,wisp} and effectively leveraging
the relatively high efficiency of SRAM is a key challenge for task-based
intermittent systems.

>>>>

Advances in processor efficiency along with the development of energy-harvesting power systems has created a new category of devices that require neither a battery nor a tethered power supply~\cite{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}. These devices operate entirely using energy extracted from their environment, such as RF energy~\cite{rf_powered_computing_gollakota_2014}, photovoltaic components~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}. Incorporating compute, storage, sensing, and communication hardware, such energy-independent devices are a promising underlying technology for applications in emerging areas such as Internet of Things (IoT)~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat}.

While promising, energy-harvesting devices present a unique design challenge in
that they operate only {\em intermittently} when energy is available~\cite{hicks_isca_2017,lucia_snapl_2017}. An
energy-harvesting device buffers energy from its environment, e.g., in a
capacitor~\cite[Fig. 3]{gorlatova_tmc_2013},~\cite[Fig. 1]{gunduz_commag_2014}. When a threshold quantity of energy accumulates, the device begins operating. Harvestable energy sources typically produce relatively very little power compared to a platform's operating power level---compare for instance~\cite[Table III and V]{prasad_comst_2014} with~\cite[Table I]{carrano_cst_2014}. A devices operates for a brief period of time and when buffered energy is depleted, shuts down and begins recharging to later operate again. As an example, inter-task time may even exceed 20\,s for wireless energy-harvesting battery-less vivo digestive tract monitoring system~\cite[Fig. 3c]{nadeau_naturebio_2017}. Charge and discharge times vary by device and some of them~\cite{wisp} may fail $\approx$10 to $\approx$100 times per second~\cite[Fig. 1]{tan_infocom_2016},~\cite[Fig. 1]{mementos}.

\todo{Put exact reference for "tictpl"}{Brandon}

Software in an energy-harvesting system operates according to the {\em
intermittent execution model}~\cite[Sec. 2]{dino}~\cite[Fig. 2]{lucia_snapl_2017}, with bursts of execution that are interrupted by failure periods. When power fails, a device loses its volatile state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e., FRAM, Flash. While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\cite{mspcdino,edb}.

%FRAM costs more than FRAM -- need to figure out how to execute out of SRAM

New programming and execution models~\cite{dino,ratchet,chain,alpaca} and
architecture mechanisms~\cite{hicks_isca_2017,idetic,nvp} help avoid data
inconsistency. New architectures require costly hardware changes and are
inapplicable to today's systems. New programming models and compilers force
programmers to use new memory models~\cite{chain,ratchet}, which may inhibit
adoption. New memory models may also increase the amount of non-volatile
memory accesses by requiring non-volatile versioning~\cite{chain} or completely
precluding the use of volatile memory~\cite{ratchet}. Increasing pressure on
non-volatile memory or precluding the use of volatile memory may degrade
efficiency; e.g., FRAM has a higher access energy and latency than
SRAM~\cite{alpaca,wisp}. Commercially available devices are often provisioned
with much less SRAM than FRAM~\cite{wolverine,wisp} and effectively leveraging
the relatively high efficiency of SRAM is a key challenge for task-based
intermittent systems.
